---
title: "r-postgis"
author: "Jay Matsushiba"
date: "2024-03-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sf)
library(RPostgres) 
library(tidyverse)
```

## R and PostGIS
This is sample code for accessing PostGIS database, running on a local Docker container. 

Following documentation for sf: https://r-spatial.github.io/sf/articles/sf2.html 

### Using st_read
In this case, the data requested from the db is read entirey into R. This would work well for smaller tables, but not for extremely large datasets.  
```{r read}
st_layers("postgresql://postgres:fr24Password@localhost:5432/shark")
# multiple different layers or tables to read within the shark db

# reads the shark table in the shark db
sharks <- st_read("postgresql://postgres:fr24Password@localhost:5432/shark", "shark")

ggplot() +
  geom_sf(data = sharks)
```

### reading and writing directly to and from spatial databases
Using this tutorial for understanding how to connect to PostgreSQL database 
https://www.r-bloggers.com/2020/08/a-crash-course-on-postgresql-for-r-users-2/ 

For some reason, RPostgreSQL package caused RStudio to crash. Using the RPostgres package works fine. 
Tried this tutorial first, then ran into that problem: 
https://www.r-bloggers.com/2015/05/getting-started-with-postgresql-in-r/ 

```{r}
# postgresql://postgres:fr24Password@localhost:5432/shark
conn = DBI::dbConnect(Postgres(),
                 dbname = "shark",
                 host = "localhost",
                 port = "5432",
                 user = "postgres",
                 password = "fr24Password"
                 )
dbExistsTable(conn, "shark") # TRUE

# get entire layer 
shark = st_read(conn, "shark")

# query for first species in layer
shark_1 = st_read(conn, query = "SELECT * FROM shark LIMIT 1")

dbDisconnect(conn)
```

